Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-04-30 paper/graph.tex
Minor typo
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@595 3fc257bb-8550-0410-81b3-a5c103da83f7
Added SSSP algorithm
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@596 3fc257bb-8550-0410-81b3-a5c103da83f7

\> 2 Collate: \> Edge as key \\
\> 2 Reduce: \> Emit zone re-assignments \\
       \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
       \> \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
       \> \> \> if $Z_i$ and $Z_j$ are different: \\
      \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
\> 2 Exit: \> if no output by Reduce 2 \\
\> 3 Map: \> Invert vertex/zone pairs \\
    \> \> \> input Key = $V_i$, Value = $Z_i$ \\
    \> \> \> if $Z_i$ is not partitioned: \\
   \> \> \> \> output Key = $Z_i$, Value = $V_i$ \\
    \> \> \> else: \\
   \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
\> 3 Add: \> Changed zones ($Z_i$,$Z_{winner}$) \\
    \> \> \> if $Z_i$ is not partitioned: \\
   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
    \> \> \> else: \\
   \> \> \> \> output Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
\> 3 Collate: \> Zone ID as key \\
\> 3 Reduce: \> Emit new zone assignment of each vertex \\
       \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
       \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
       \> \> \> partition $Z_{new}$ if number of $V$ > threshhold \\
       \> \> \> for each $V$ in MultiValue: \\
      \> \> \> \> output Key = $V_i$, Value = $Z_{new}$

@@

Discussion of SSSP.

Discuss new optimization in SSSP?

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output: Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

@@

>>>>

\> 2 Collate: \> Edge as key \\
\> 2 Reduce: \> Emit zone re-assignments \\
       \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
       \> \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
       \> \> \> if $Z_i$ and $Z_j$ are different: \\
      \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
\> 2 Exit: \> if no output by Reduce 2 \\
\> 3 Map: \> Invert vertex/zone pairs \\
    \> \> \> input Key = $V_i$, Value = $Z_i$ \\
    \> \> \> if $Z_i$ is not partitioned: \\
   \> \> \> \> output Key = $Z_i$, Value = $V_i$ \\
    \> \> \> else: \\
   \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
\> 3 Add: \> Changed zones ($Z_i$,$Z_{winner}$) \\
    \> \> \> if $Z_i$ is not partitioned: \\
   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
    \> \> \> else: \\
   \> \> \> \> output Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
\> 3 Collate: \> Zone ID as key \\
\> 3 Reduce: \> Emit new zone assignment of each vertex \\
       \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
       \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
       \> \> \> partition $Z_{new}$ if number of $V >$ threshhold \\
       \> \> \> for each $V$ in MultiValue: \\
      \> \> \> \> output Key = $V_i$, Value = $Z_{new}$

@@

The single-source shortest path (SSSP) algorithm computes, for a directed graph
with weighted edges,
the shortest weighted distance from a given source vertex to
all other vertices in the graph.  This operation is straightforward
when global graph information is available.  The source vertex is labeled with
distance zero.  Then in each iteration, edge adjacencies of labeled vertices
are followed, and adjacent vertices are labeled with updated distances.  When
no distances change, the iterations are complete.  Most importantly, only
edges from modified vertices are visited in each iteration.

In MapReduce, global information about edge adjacencies and labeled vertices
is not available.  Instead, following the Bellman-Ford-style 
algorithm of~\cite{SSSPMapReduce, BellmanFord},
every edge of the graph is visited in each iteration,
although labels for an edge's vertices are updated only if the breadth-first
traversal has reached those vertices.  The algorithm begins with a {\it map()}
function loading edges $(v_i, v_j)$ and source vertex $v_s$ with distance 0.
In each iteration, the edges $(v_i, v_j)$ are collated with respect to
$v_i$, and a {\it reduce()} operation updates each $v_j$'s distance if
$v_i$'s distance is non-zero.  The algorithm is fully described in
Figure~\ref{fig:sssp}.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxx\=xxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

Input:  \\
\> Source vertex $v_s$.\\
\> MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$ (initially $\infty$):  \\
\> \> Key = $v_i$; Value = $\infty$.\\
\> MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
Map $V$:  Set distance for source vertex $v_s$ to zero. \\
\> Input:  Key = $v_i$; Value = $\infty$ \\
\> Output:  Key = $v_i$; Value = 0 if $v_i = v_s$, otherwise Value = $\infty$. \\
while not done: \\
\> done = 1; \\
\> Add $E$ to $V$. \\
\> Collate $V$:  Key is vertex $v_i$. \\
\> Reduce $V$:  Loop over all edges to perform breadth-first search from labeled vertices. \\
\> \> Input:  Key = $v_i$; Multivalue = edges $[v_j, w_{ij}]$, candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output:  Key = $v_i$; Value = $d_{min}$ (updated distance for $v_i$) \\
\> \> \> and if $d_{min} \ne d_i$, Key = $v_j$; Value = $d_{min} + w_{ij}$ for each edge ($v_i, v_j$). \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.

@@

As with PageRank and R-MAT generation, 
significant savings in communication and execution time
can be achieved by pre-aggregating certain MapReduce objects and 
storing data in more than one MapReduce object.  The improved algorithm
is shown in Figure~\ref{fig:sssp2}.
